---
title: "Barcode_analysis"
author: "Dawn Lin"
date: "12/01/2020"
output: html_document
---

# set directory
```{r}
setwd("~/Desktop/FL_DC_paper/Barcoding/Barcode_data")
```

# reading in raw count data
```{r cars}
DL191 = read.table("DL191_noE_pool.txt",header=TRUE)
DL198 = read.table("DL198_noE_pool.txt",header=TRUE)
DL33 = read.table("DL33_pool.txt",header=TRUE)
DL70 = read.table("DL70_pool.txt",header=TRUE)
DL72 = read.table("DL72_pool.txt",header=TRUE)

dt=rbind(DL191,DL198,DL33,DL70,DL72)

```

# load packages
```{r pressure, echo=FALSE}
library(Rtsne)
library(ggplot2)
library(grid)
library(reshape2)
library(ggdendro)
library(ggbeeswarm)
library(dplyr)
library(gridExtra)
library(gplots)
library(pheatmap)
library(tidyverse)
```

# Specify color
```{r}
library(RColorBrewer)
colors = rev(brewer.pal(7,"RdYlBu"))
my_colour = c("#09A2D1","#F2AA4CFF")
```

# tsne analysis
```{r}
colnames(dt)
set.seed(1234)
tsne=Rtsne(dt[,c("cDC1","cDC2","pDC","B","mye")],perplexity=40,check_duplicates=FALSE) 

# color code on cell type
cDC1=
  ggplot(data = NULL, aes(x=tsne$Y[,1],y=tsne$Y[,2],col=dt[,"cDC1"]))+
  geom_point(size=0.5)+
  scale_color_gradientn(colours=colors)+
  theme_classic()+
  labs(title="cDC1",x="",y="") 
cDC2=
  ggplot(data = NULL, aes(x=tsne$Y[,1],y=tsne$Y[,2],col=dt[,"cDC2"]))+
  geom_point(size=0.5)+
  scale_color_gradientn(colours=colors)+
  theme_classic()+
    labs(title="cDC2",x="",y="")
pDC=
  ggplot(data = NULL, aes(x=tsne$Y[,1],y=tsne$Y[,2],col=dt[,"pDC"]))+
  geom_point(size=0.5)+
  scale_color_gradientn(colours=colors)+
  theme_classic()+
    labs(title="pDC",x="",y="") 

mye=
  ggplot(data = NULL, aes(x=tsne$Y[,1],y=tsne$Y[,2],col=dt[,"mye"]))+
  geom_point(size=0.5)+
  scale_color_gradientn(colours=colors)+
  theme_classic()+
    labs(title="M",x="",y="") 
B=
  ggplot(data = NULL, aes(x=tsne$Y[,1],y=tsne$Y[,2],col=dt[,"B"]))+
  geom_point(size=0.5)+
  scale_color_gradientn(colours=colors)+
  theme_classic()+
    labs(title="B",x="",y="") 

# color code on treatment
D = data.frame(x = tsne$Y[,1], y = tsne$Y[,2], treatment = dt$`treatment`)
D$treatment <- factor(D$treatment,levels = c("PBS", "FL"))
treatment=
  ggplot(D)+
  geom_point(aes(x,y,colour = factor(treatment)),alpha=1,size=1,shape=21)+
  scale_colour_manual(values = my_colour)+
  theme_classic()+
  ggtitle("treatment")

D = data.frame(x = tsne$Y[,1], y = tsne$Y[,2], exp = dt$`exp`)
exp=
  ggplot(D)+
  geom_point(aes(x,y,colour = factor(exp)),alpha=1,size=1,shape=21)+
  theme_classic()+
  ggtitle("expID")
```

## plot tsne
```{r}
pdf("tsne.pdf", height = 9,width = 18)
multiplot(cDC1,cDC2,pDC,mye,B,treatment, cols=3)
dev.off()
```

## add t-SNE coordinate to dataset
```{r}
D = data.frame(x = tsne$Y[,1], y = tsne$Y[,2])
df = cbind(D,dt)
```


# DBSCAN clustering
```{r}
library(fpc)
library(dbscan)
library(factoextra)

set.seed(100)
# eps: the radius of neighborhood around a point x
# MinPts: minimum number of neighbors within “eps” radius
db=fpc::dbscan(tsne$Y, eps=2.2, MinPts=20)
print(db)
```

## plot DBSCAN on tsne
```{r}
fviz_cluster(db, data=tsne$Y,stand = FALSE,
             ellipse = FALSE, 
             show.clust.cent = FALSE,
             aes=c(tsne$Y[,1],tsne$Y[,2]),
             xlab = FALSE, ylab = FALSE,
             geom = "point", shape=16, pointsize=1, 
             ggtheme = theme_classic())
```

# Add Cluster ID to dataframe
```{r}
df$cluster=db$cluster
write.table(df, "pool_tsne_s1234p40_db_e22m20.txt", sep="\t", row.names=FALSE, col.names=TRUE)
```

# reorder dataframe
```{r}
df = df[c("cDC1","cDC2","pDC","mye","B","eos","mon","neu",
         "treatment","exp","mouse","cluster","x","y")]

df$cluster <- sub("^", "cluster ", df$cluster)

D = data.frame()
cluster = c("cluster 1",
            "cluster 11",
            "cluster 18",
            "cluster 15",
            "cluster 19",
            "cluster 16",
            "cluster 20",
            "cluster 17",
            "cluster 9",
            "cluster 4",
            "cluster 12",
            "cluster 7",
            "cluster 6",
            "cluster 10",
            "cluster 2",
            "cluster 14",
            "cluster 13",
            "cluster 8",
            "cluster 5",
            "cluster 3")
for (c in cluster)
{ 
  dtc=subset(df, df$cluster==c)
  dtc=dtc[order(rowSums(dtc[,c(1:5)]),decreasing=T),]
  D=rbind(D,dtc)
}

df_order = D

```

# plot heatmap
```{r}
anno_row = df_order[c(9,12)]

my_colour = list(
  treatment = c(PBS = "#09A2D1", FL = "#F2AA4CFF")
)

pheatmap(data.matrix(df_order[,c(1:5)]),
         cluster_rows = F, cluster_cols = T, 
         annotation_row = anno_row,
         annotation_colors = my_colour,
         filename = "heatmap.pdf",
         width = 4,
         height = 8
         )
```

# Count #bc per cluster from each mouse
```{r}
D=data.frame()
cluster=unique(df$cluster)
for (c in cluster)
{ 
  dtc=subset(df, df$cluster==c)
  mouse=unique(dt$mouse)
  a=data.frame()
  for (m in mouse)
  {
    counts=sum(dtc$mouse==m) 
    b=data.frame(mouse=m,counts=counts)
    b$cluster=c
    a=rbind(a,b)
  }
  D=rbind(D,a)
}
write.table(D, "cluster_mouse_bc.txt", sep="\t", row.names=FALSE, col.names=TRUE)
```


# calculate proportional contribution to cDC1s by each clusters
```{r}
a = as.data.frame(sinh(df[1:8]))# transform read counts back to linear
b = as.data.frame(df[9:14])
dt.n = cbind(a,b)

cluster=unique(dt.n$cluster)
D=data.frame()
for (c in cluster)
{ 
  if(c==0)next;
  dtc=subset(dt.n, dt.n$cluster==c)
  
  mouse=unique(dtc$mouse)
  a=data.frame()
  for (m in mouse)
  {
    dtm=subset(dt.n, dt.n$mouse==m)
    dtcm=subset(dtc, dtc$mouse==m)
    cDC1=sum(dtcm["cDC1"])/sum(dtm["cDC1"])
    cDC2=sum(dtcm["cDC2"])/sum(dtm["cDC2"])
    pDC=sum(dtcm["pDC"])/sum(dtm["pDC"])
    mye=sum(dtcm["mye"])/sum(dtm["mye"])
    B=sum(dtcm["B"])/sum(dtm["B"])
    treatment=unique(dtcm$treatment)
    b=data.frame(clsuter=c,mouse=m,treatment=treatment,
                 cDC1=cDC1,cDC2=cDC2,pDC=pDC,mye=mye)
    a=rbind(a,b)
  }
  D=rbind(D,a)
}

write.table(D, "proportional contribution to pop.txt", sep="\t", row.names=FALSE, col.names=TRUE)

```

# normalize data to cell number
```{r}
bc.PBS = subset(dt.n, dt.n$treatment=="PBS")

bc.PBS$cDC1 = bc.PBS$cDC1/1.4e7*53257 # total #cDC1 produced by 14 PBS-treated mice
  bc.PBS$cDC2 = bc.PBS$cDC2/1.4e7*44323
  bc.PBS$pDC = bc.PBS$pDC/1.4e7*36509
  bc.PBS$mye = bc.PBS$mye/1.4e7*823318
  bc.PBS$B = bc.PBS$B/1.4e7*869332
  bc.PBS$eos = bc.PBS$eos/1.4e7*39623
  bc.PBS$mon = bc.PBS$mon/1.4e7*341084
  bc.PBS$neu = bc.PBS$neu/1.4e7*442610

bc.FL = subset(dt.n, dt.n$treatment=="FL")
  bc.FL$cDC1 = bc.FL$cDC1/1.6e7*1082433 #total #cDC1 produced by 14 FL-treated mice
  bc.FL$cDC2 = bc.FL$cDC2/1.6e7*224557 
  bc.FL$pDC = bc.FL$pDC/1.6e7*102281 
  bc.FL$mye = bc.FL$mye/1.6e7*1174073 
  bc.FL$B = bc.FL$B/1.6e7*2270197 
  bc.FL$eos = bc.FL$eos/1.6e7*46699 
  bc.FL$mon = bc.FL$mon/1.6e7*509426 
  bc.FL$neu = bc.FL$neu/1.6e7*617947 

dt.n = rbind(bc.PBS,bc.FL)
write.table(dt.n, "pool_cellnumber.txt", sep="\t", row.names=FALSE, col.names=TRUE)

```

# perform stat analysis on clone size between PBS vs FL per cluster
```{r}
D.pvalue=data.frame()
D.foldchange=data.frame()
cluster=unique(dt.n$cluster)
for (c in cluster)
{ 
  if(c==0)next;
  cnc=subset(dt.n, dt.n$cluster==c)
  cnc.fl=subset(cnc, cnc$treatment=="FL")
  cnc.pbs=subset(cnc, cnc$treatment=="PBS")
  # count bc number
  n.fl=nrow(cnc.fl)
  n.pbs=nrow(cnc.pbs)
  # stat test in clone size per cell type
  cDC1=t.test(cnc.fl$cDC1, cnc.pbs$cDC1)
   p.cDC1=cDC1$p.value
  cDC2=t.test(cnc.fl$cDC2, cnc.pbs$cDC2)
   p.cDC2=cDC2$p.value
  pDC=t.test(cnc.fl$pDC, cnc.pbs$pDC)
   p.pDC=pDC$p.value
  mye=t.test(cnc.fl$mye, cnc.pbs$mye)
   p.mye=mye$p.value
  B=t.test(cnc.fl$B, cnc.pbs$B)
   p.B=B$p.value
  # calculate average fold change in clone size
   fc.cDC1=colMeans(cnc.fl["cDC1"])/colMeans(cnc.pbs["cDC1"])
   fc.cDC2=colMeans(cnc.fl["cDC2"])/colMeans(cnc.pbs["cDC2"])
   fc.pDC=colMeans(cnc.fl["pDC"])/colMeans(cnc.pbs["pDC"])
   fc.mye=colMeans(cnc.fl["mye"])/colMeans(cnc.pbs["mye"])
   fc.B=colMeans(cnc.fl["B"])/colMeans(cnc.pbs["B"])
    
   p.Sum=data.frame(cluster=c,cDC1=p.cDC1, cDC2=p.cDC2, pDC=p.pDC,mye=p.mye, B=p.B)
   fc.Sum=data.frame(cluster=c,cDC1=fc.cDC1, cDC2=fc.cDC2, pDC=fc.pDC,mye=fc.mye, B=fc.B)

   D.pvalue=rbind(D.pvalue,p.Sum)
   D.foldchange=rbind(D.foldchange,fc.Sum)
}

write.table(D.pvalue, "clone size_p value.txt", sep="\t", row.names=FALSE, col.names=TRUE)
write.table(D.foldchange,  "clone size_fold change.txt",  sep="\t", row.names=FALSE, col.names=TRUE)
```

# Violin plots showing distribution of clone size
```{r}
cluster=unique(dt.n$cluster)
for (c in cluster)
{ 
  if(c==0)next;
  cell.num.c=subset(dt.n, dt.n$cluster==c)
  cell.num.c$treatment <- factor(cell.num.c$treatment,levels = c("PBS", "FL"))
  
  P1=
    ggplot(data=cell.num.c)+
    geom_violin(aes(x=factor(treatment),y=cDC1, fill=factor(treatment)))+
    scale_y_log10(limit=c(1,100000))+
    scale_fill_manual(values = my_colour)+
    theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())
  P2=
    ggplot(data=cell.num.c)+
    geom_violin(aes(x=factor(treatment),y=cDC2,fill=factor(treatment)))+
    scale_y_log10(limit=c(1,100000))+
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
    scale_fill_manual(values = my_colour)+
    theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())
  P3=ggplot(data=cell.num.c)+
    geom_violin(aes(x=factor(treatment),y=pDC,fill=factor(treatment)))+
    scale_y_log10(limit=c(1,100000))+
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
    scale_fill_manual(values = my_colour)+
    theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())
  P4=ggplot(data=cell.num.c)+
    geom_violin(aes(x=factor(treatment),y=mye,fill=factor(treatment)))+
    scale_y_log10(limit=c(1,100000))+
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
    scale_fill_manual(values = my_colour)+
    theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())
  P5=ggplot(data=cell.num.c)+
    geom_violin(aes(x=factor(treatment),y=B,fill=factor(treatment)))+
    scale_y_log10(limit=c(1,100000))+
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
    scale_fill_manual(values = my_colour)+
    theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())
  
  pdf(file=paste("cluster", c, "clone size.pdf"),heigh =1,width=4)
  multiplot(P1,P2,P3,P4,P5, cols=5)
  dev.off()
}
```
